---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sve

The goal of sve is to facilitate the calculation of symmetric vaccine efficacy (SVE) and corresponding variance and confidence intervals.

## Installation

You can install the development version of sve like so:

``` r
devtools::install_github("LucyMcGowan/sve")
```

## Example

```{r}
library(sve)
```

The core function, `est_sve()`, computes the symmetric vaccine efficacy and its confidence interval given the event proportions and sample sizes for the vaccinated and unvaccinated groups.

**Example:**

* unvaccinated 10% infection rate  
* vaccinated 5% infection rate  
* $n_0=n_1=1,000$

```{r}
est_sve(p0 = 0.10, 
        p1 = 0.05, 
        n0 = 1000, 
        n1 = 1000)
```

By default, confidence intervals are calculated on a transformed scale and then back-transformed in order to ensure that they remain between -1 and 1. If you do not want to use this transformation, you can set the option `transform = FALSE`:

```{r}
est_sve(p0 = 0.10, 
        p1 = 0.05, 
        n0 = 1000, 
        n1 = 1000,
        transform = FALSE)
```
## Methods overview

Method overview
The symmetric vaccine efficacy (SVE) is defined as

$$
\text{SVE}=\frac{2(p_0-p_1)}{p_0+p_1+|p_0-p_1|},
$$
 
where:
* $p_0$ = event proportion in the unvaccinated group  
* $p_1$  = event proportion in the vaccinated group

This formulation ensures the estimator is bounded between -1 and 1.

